%---------------------------------------------------------------
\section{Research Questions}\label{research-questions}
List the specific research questions you are addressing from the MiniChallenge. For each question, include a short description of how you plan to approach it.

\begin{quote}
    FishEye analysts need your help to perform geographic and temporal analysis of the CatchNet data so they can prevent illegal fishing from happening again. Your task is to develop new visual analytics tools and workflows that can be used to discover and understand signatures of different types of behavior. Can you use your tool to visualize a signature of SouthSeafood Express Corp’s illegal behavior? FishEye needs your help to develop a workflow to find other instances of illegal behavior.

\begin{enumerate}
    \item FishEye analysts have long wanted to better understand the flow of commercially caught fish through Oceanus’s many ports. But as they were loading data into CatchNet, they discovered they had purchased the wrong port records. They wanted to get the ship off-load records, but they instead got the port-exit records (essentially trucks/trains leaving the port area). Port exit records do not include which vessel that delivered the products. Given this limitation, develop a visualization system to associate vessels with their probable cargos. Which vessels deliver which products and when? What are the seasonal trends and anomalies in the port exit records?\\
    \item Develop visualizations that illustrate the inappropriate behavior of SouthSeafood Express Corp vessels. How do their movement and catch contents compare to other fishing vessels? When and where did SouthSeafood Express Corp vessels perform their illegal fishing? How many different types of suspicious behaviors are observed? Use visual evidence to justify your conclusions.\\
    \item To support further Fisheye investigations, develop visual analytics workflows that allow you to discover other vessels engaging in behaviors similar to SouthSeafood Express Corp’s illegal activities? Provide visual evidence of the similarities.\\
    \item How did fishing activity change after SouthSeafood Express Corp was caught? What new behaviors in the Oceanus commercial fishing community are most suspicious and why?
\end{enumerate}
    Note: the VAST challenge is focused on visual analytics and graphical figures should be included with your response to each question. Please include a reasonable number of figures for each question (no more than about 6) and keep written responses as brief as possible (around 250 words per question). Participants are encouraged to new visual representations rather than relying on traditional or existing approaches.
\end{quote}